---
title: "Demo"
author: ""
date: "19/07/2020"
output:
  html_document:
    toc: true
    toc_float: true
---



## Data

### dat


* id (string)
* pet (string)
  * Levels
    * dog
    * cat
  * Ordered: FALSE
* time (string)
  * Levels
    * morning
    * night
  * Ordered: FALSE
* y (float)


```{r}
dat <- read.csv('data/dat_data.tsv', sep='\t')
```



## Analysis 1: A1 {#analysis_1}

```{r}
    scienceverse::t_test(y ~ pet, dat)
```

### Stored Results

* statistic: 4.1843
* parameter: 117.806
* p.value: 1e-04
* conf.int: 
  * 2.1027
  * 5.8814
* estimate: 
  * 11.0571
  * 7.065
* null.value: 0
* stderr: 0.954
* alternative: two.sided
* method: Welch Two Sample t-test
* data.name: y by pet
* n: 
  * 60
  * 60

## Analysis 2: A2 {#analysis_2}

```{r}
    scienceverse::t_test(y ~ time, dat)
```

### Stored Results

* statistic: -0.045
* parameter: 117.9801
* p.value: 0.9641
* conf.int: 
  * -2.0706
  * 1.9785
* estimate: 
  * 9.038
  * 9.0841
* null.value: 0
* stderr: 1.0224
* alternative: two.sided
* method: Welch Two Sample t-test
* data.name: y by time
* n: 
  * 60
  * 60


