---
title: "Demo"
author: ""
date: "19/07/2020"
output:
  html_document:
    toc: true
    toc_float: true
---



## Data

### dat


* id (string)
* group (string)
  * Levels
    * A
    * B
  * Ordered: FALSE
* time (string)
  * Levels
    * morning
    * night
  * Ordered: FALSE
* size (float)


```{r}
dat <- read.csv('data/dat_data.tsv', sep='\t')
```



## Analysis 1: A1 {#analysis_1}

```{r}
    t.test(size ~ group, dat)
```

### Stored Results

* statistic: -0.2403
* parameter: 111.5102
* p.value: 0.8106
* conf.int: 
  * -4.1552
  * 3.2564
* estimate: 
  * 101.1039
  * 101.5532
* null.value: 0
* stderr: 1.8702
* alternative: two.sided
* method: Welch Two Sample t-test
* data.name: size by group

## Analysis 2: A2 {#analysis_2}

```{r}
    t.test(size ~ time, dat)
```

### Stored Results

* statistic: 2.2828
* parameter: 116.126
* p.value: 0.0243
* conf.int: 
  * 0.5532
  * 7.8049
* estimate: 
  * 103.4181
  * 99.239
* null.value: 0
* stderr: 1.8307
* alternative: two.sided
* method: Welch Two Sample t-test
* data.name: size by time


