---
title: "Demo"
author: ""
date: "30/09/2020"
output:
  html_document:
    toc: true
    toc_float: true
---



## Data

### dat


* id (string)
* pet (string)
  * Levels
    * dog
    * cat
  * Ordered: FALSE
* time (string)
  * Levels
    * morning
    * night
  * Ordered: FALSE
* y (float): Happiness Score


```{r}
dat <- read.csv('data/dat_data.tsv', sep='\t')
```



## Analysis 1: A1 {#analysis_1}

```{r}
    scienceverse::t_test(y ~ pet, dat)
```

### Stored Results

* statistic: `3.0784`
* parameter: `117.757`
* p.value: `0.0026`
* conf.int: 
  * `0.9353`
  * `4.3087`
* estimate: 
  * `10.7868`
  * `8.1647`
* null.value: `0`
* stderr: `0.8517`
* alternative: `two.sided`
* method: `Welch Two Sample t-test`
* data.name: `y by pet`
* n: 
  * `60`
  * `60`

## Analysis 2: A2 {#analysis_2}

```{r}
    scienceverse::t_test(y ~ time, dat)
```

### Stored Results

* statistic: `-0.8188`
* parameter: `117.9216`
* p.value: `0.4145`
* conf.int: 
  * `-2.471`
  * `1.0253`
* estimate: 
  * `9.1143`
  * `9.8372`
* null.value: `0`
* stderr: `0.8828`
* alternative: `two.sided`
* method: `Welch Two Sample t-test`
* data.name: `y by time`
* n: 
  * `60`
  * `60`


