---
title: "Demo"
author: ""
date: "17/10/2020"
output:
  html_document:
    toc: true
    toc_float: true
---



## Data

### dat


* id (string)
* pet (string)
  * Levels
    * dog
    * cat
  * Ordered: FALSE
* time (string)
  * Levels
    * morning
    * night
  * Ordered: FALSE
* y (float): Happiness Score


```{r}
dat <- read.csv('data/dat_data.tsv', sep='\t')
```



## Analysis 1: A1 {#analysis_1}

```{r}
scienceverse::t_test(y ~ pet, dat)
```

### Stored Results

* statistic: `2.5266`
* parameter: `117.118`
* p.value: `0.0129`
* conf.int: 
    1. `0.4967`
    2. `4.0989`
* estimate: 
    1. `9.0335`
    2. `6.7357`
* null.value: `0`
* stderr: `0.9094`
* alternative: `two.sided`
* method: `Welch Two Sample t-test`
* data.name: `y by pet`
* n: 
    1. `60`
    2. `60`

## Analysis 2: A2 {#analysis_2}

```{r}
scienceverse::t_test(y ~ time, dat)
```

### Stored Results

* statistic: `0.105`
* parameter: `117.5104`
* p.value: `0.9166`
* conf.int: 
    1. `-1.751`
    2. `1.947`
* estimate: 
    1. `7.9336`
    2. `7.8356`
* null.value: `0`
* stderr: `0.9337`
* alternative: `two.sided`
* method: `Welch Two Sample t-test`
* data.name: `y by time`
* n: 
    1. `60`
    2. `60`


