---
title: "Codebook Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Codebook Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(scienceverse)
```

Codebooks in scienceverse follow the Psych-DS 0.1.0 format by default as set up in the function `faux::codebook()`, although you can manually add your own codebook to the JSON file.

## Automatic Codebooks for Simulated Data

When you add simulated data to a study object, the codebook is generated by default.

```{r}
study <- study("Demo") %>%
  add_sim_data("pets",
               between = list(pet = c(cat = "Cat Owners", 
                                      dog = "Dog Owners")),
               n = c(4, 6),
               dv = list(happy = "Happiness Score"),
               id = list(id = "Subject ID"),
               mu = c(10, 12),
               sd = 4)
```

You can retrieve the codebook with the function `get_codebook()`. If you don't specify the data_id, it defaults to the first data set in the study.

```{r}
cb <- get_codebook(study, data_id = "pets")
```

If you just type the codebook object into the console, you'll see the info formatted like this.

```{r}
cb
```

But the codebook is actually a nested list formatted like this:

```{r}
str(cb)
```

You can also view it in JSON format.

```{r}
get_codebook(study, data_id = "pets", as_json = TRUE)
```

## Run Codebook on Existing Data

You can run the codebook function on existing data, but will need to manually input some things. For now, the supported properties are: "description", "privacy", "type", "propertyID", "minValue", "maxValue", "levels", "ordered", "na", "naValues", "alternateName", "unitCode". You can set unsupported properties, but this will produce a warning.

```{r}


vardesc <- list(
  description = list(
    mpg = "Miles/(US) gallon",
    cyl = "Number of cylinders",
    disp = "Displacement (cu.in.)",
    hp = "Gross horsepower",
    drat = "Rear axle ratio",
    wt = "Weight (1000 lbs)",
    qsec = "1/4 mile time",
    vs = "Engine (0 = V-shaped, 1 = straight)",
    am = "Transmission (0 = automatic, 1 = manual)",
    gear = "Number of forward gears",
    carb = "Number of carburetors"
  ),
  # min and max values can be set manually or from data
  # min and max are often outside the observed range
  minValue = list(mpg = 0, cyl = min(mtcars$cyl)),
  maxValue = list(cyl = max(mtcars$cyl)),
  levels = list(
    vs = c("0" = "V-shaped", "1" = "straight"),
    am = c("0" = "automatic", "1" = "manual")
  ),
  type = list(
    cyl = "integer",
    hp = "integer",
    vs = "factor",
    am = "factor",
    gear = "integer",
    carb = "integer"
  )
)

codebook(mtcars, "Motor Trend Car Road Tests", vardesc)
```

