---
title: "Crediting Authors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Crediting Authors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(scienceverse)
```

## CRediT

Scienceverse supports the [CRediT taxonomy](https://casrai.org/credit/) for authorship roles.

Use the function `credit_roles()` to view a list of possible roles and their abbreviations. Set `display` to `"names"` or `"abbr"` to just list the names or abbreviations.

```{r}
credit_roles()
```


```{r}
credit_roles(display = "names")
```

```{r}
credit_roles(display = "abbr")
```

## ORCiD

Scienceverse supports recording and searching for [ORCiDs](https://orcid.org/). Every researcher should have an ORCiD to link their papers and avoid problems with name changes.

You can use the `get_orcid()` function to search for an ORCiD quickly, although you might need to go to their website for more advanced search if they have a common name and you get multiple hits. This function requires a web connection.

```{r}
ldb_orcid <- get_orcid("DeBruine", "Lisa")
```

## Adding an author to a study

You can add authors to a study object with the `add_author()` function. Authors must have a `surname`, and you can optionally add a `given` name, `orcid` and `roles` (using the names, abbreviations, or indices above). You can also make up any other category of information, such as `email` or `affiliation`.

```{r, results='asis'}
# set up an empty study
s <- study("Demo") 

s <- add_author(s, 
                surname = "DeBruine",
                given = "Lisa M.",
                orcid = ldb_orcid, 
                roles = c("con", "sof", "edi"),
                email = "debruine@gmail.com") %>%
  add_author(surname = "Lakens",
             given = "Daniel",
             orcid = get_orcid("Lakens", "Daniel"),
             roles = c(1, 13, 14))

s$authors
```



## CRediT format

You can use the `author_jats()` function to format a study's author list in the machine-readable JATS format requested by CRediT.

```{r}
jats <- author_jats(s)
```

The output of `author_jats()` looks like this in the source code of a webpage:

<pre>
`r htmltools::htmlEscape(jats)`
</pre>

 







