---
title: "Quick Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

```{r setup}
library(scienceverse)
```

One goal of scienceverse is to automate the evaluation of predictions. A researcher specifies the prediction in the preregistration, collects the data, and the scienceverse R package (or other platforms that handle scienceverse-formatted files) can then use the preregistration file and the data to automatically evaluate whether the predictions were confirmed or not.

The demo below follows the example from: 

Lakens, D., & DeBruine, L. M. (2020, January 27). Improving Transparency, Falsifiability, and Rigour by Making Hypothesis Tests Machine Readable. <https://doi.org/10.31234/osf.io/5xcda>

## Set up the study

Set up the study with a name and any additional info you want to add.

```{r}
study <- study(name = "Kinship and Prosocial Behaviour",
               abstract = "A reanalysis of data from DeBruine (2002) Facial Resemblance Enhances Trust, PRSLB.")
```

## Authors

Add each author in the following format. Use the function `credit_roles()` to see a list of the roles and their descriptions. Use `credit_roles("name")` or `credit_roles("abbr")` to just see their names or abbreviations. You can also add any further info, like an orcid or email address.

```{r}
study <- add_author(study, 
                    orcid = "0000-0002-7523-5539",
                    surname = "DeBruine",
                    given = "Lisa M.",
                    roles = c("con", "dat", "sof", "dra", "edi"),
                    email = "lisa.debruine@glasgow.ac.uk") %>%
         add_author(orcid = "0000-0002-0247-239X",
                    surname = "Lakens",
                    given = "Daniël",
                    roles = c("con", "ana", "dra", "edi"))
```

## Hypotheses

Now add a hypothesis with a verbal description. You can add more than one hypothesis, but this demo study only has one.

```{r}
study <- add_hypothesis(study, id = "self_pref",
                        description = "Cues of kinship will increase prosocial behaviour. Cues of kinship will be manipulated by morphed facial self-resemblance. Prosocial behaviour will be measured by responses in the trust game. The prediction is that the number of trusting AND/OR reciprocating moves will be greater to self morphs than to other morphs.")
```

## Analyses

Add all relevant analyses for testing this hypothesis in the order that they should be run. Do any data prep first. Each analysis needs an `id` for reference in the criteria later. The `code` can be R code (wrap it in {} if you need more than one line) or the file path for a .R file. You can also add other information about your analysis, such as the software it's running on. 

```{r}
study <- add_analysis(study, id = "trust",
                      code = t.test(kin$trust_self, 
                                    kin$trust_other, 
                                    paired = TRUE, 
                                    conf.level = 0.975), 
                      software = R.version.string) %>%
         add_analysis(id = "recip",
                      code = t.test(kin$recip_self,
                                    kin$recip_other,
                                    paired = TRUE,
                                    conf.level = 0.975), 
                      software = R.version.string)
```

If your analysis was run using other software or you only have summary statistics available, you can add a list of arbitrary named values like this (setting unknown values to `NA` for preregistration):

```{r, eval = FALSE}
study <- add_analysis(study, id = "tr_self",
                      code = list(
                        t = -2.50,
                        df = 23,
                        p = 0.0198,
                        ci = c(-0.9787, -0.0213),
                        estimate = -0.5,
                        stderr = 0.2,
                        description = "two-tailed paired t-test of trust_self vs recip_self"
                      ), 
                      software = "SPSS")
```


## Criteria

Next, add the criteria you will need to check to corroborate or falsify each hypothesis. Each criterion needs a unique `id` so you can reference it next in the evaluation. Then, optionally specify the `hypothesis_id` and `analysis_id` where you'll look for the result. If you omit these it will default to the last one added. For example, the analysis `trust` returns a list from the function `t.test`, which includes a value for `conf.int`. To see if the first number in this vector is larger than 0, set `result` to "conf.int[1]", `operator` to ">" , and `comparator` to $0$. The options for `operator` are ">", "<", "=", and "!=".

```{r}
study <- study %>%
  add_criterion(id = "t_lo",
                result = "conf.int[1]",
                operator = ">",
                comparator = 0.0,
                analysis_id = "trust",
                hypothesis_id = "self_pref") %>%
  add_criterion("t_hi", "conf.int[2]", ">", 0.2, "trust") %>%
  add_criterion("r_lo", "conf.int[1]", ">", 0.0, "recip") %>%
  add_criterion("r_hi", "conf.int[2]", ">", 0.2, "recip")
  
```

## Evaluation

Add evaluation criteria for corroboration and falsification. The logical `evaluation` should reference the criterion `id`s above and can consist of those ids and the symbols `(`, `)`, `&`, and `|`. You can optionally include a verbal `description` of the criterion to clarify the logical evaluation.

```{r}
study <- add_eval(study, "corroboration",
                  evaluation = "(t_lo & t_hi) | (r_lo & r_hi)",
                  description = "The hypothesis is corroborated if the 97.5% CI lower bound is greater than 0 and the 97.5% CI upper bound is greater than 0.2 (the SESOI) for either the trust or reciprocation moves.")

study <- add_eval(study, "falsification",
                  evaluation = "!t_hi & !r_hi",
                  description = "The hypothesis is falsified if the 97.5% CI upper bound is smaller than 0.2 (the SESOI) for both trust and reciprocation.")
```


## Saving

At this point, you can save your study to a machine-readable JSON-formatted file.

```{r}
study_save(study, "prereg.json", format = "json")
```

You can also export a human-readable pre-registration using the built-in "prereg" template.

```{r}
study_save(study, "prereg.html", format = "prereg")
```

## Data

The `id` you give the data can be used in the analysis code to refer to this dataset. The `data` argument can be a data frame or a file path to a data file or a PsychDS-formatted codebook. You can optionally add column parameters such as descriptions with the argument `vardesc` and other dataset parameters as additional arguments.

```{r}

kin_data <- data.frame(
  trust_self  = c(1,2,2,1,1,1,1,1,2,0,2,0,1,2,2,3,2,2,1,1,2,0,0,1),
  trust_other = c(1,2,2,0,1,0,0,0,1,0,1,0,1,1,1,0,1,2,2,0,0,0,2,1),
  recip_self  = c(0,1,3,2,1,1,1,3,3,2,3,1,1,2,3,3,3,1,1,1,3,0,3,1),
  recip_other = c(1,1,2,2,3,2,1,3,3,1,3,0,1,3,3,3,3,0,3,0,1,0,3,2)
)

desc <- list(
  description = list(
    trust_self  = "Number of trusting moves towards self-morphs",
    trust_other = "Number of trusting moves towards self-morphs",
    recip_self  = "Number of reciprocating moves towards other-morphs",
    recip_other = "Number of reciprocating moves towards other-morphs"
  ),
  dataType = rep("int", 4) # all variables are integer types
)

study <- add_data(study, id = "kin", 
                  data = kin_data, 
                  vardesc = desc, 
                  url = "https://osf.io/ewfhs/")

```

## Run Analyses

When you run `study_analyse`, the data are loaded as their `id` names and the analyses are run in order. For example, the data from the section above will be available to your analyses as a dataframe called `kin`. The analysis is run in a separate environment, so you won't see any new objects in the global environment. Therefore, changes you make to, e.g., a dataframe called `kin` in your global environment won't affect the analysis. The evaluation of each criterion and hypothesis is printed as a message.

```{r}
study <- study_analyse(study)
```


## Save and reload

At this point, you can save the post-registration version of your study to a JSON-formatted file and also export a human-readable post-registration report.

```{r}
study_save(study, "postreg.json", "json")
study_save(study, "postreg.html", "postreg")
```

You can reload and edit your study by loading the .json file with the study() function.

```{r}
study2 <- study("postreg.json")

study2 # displays a summary
```




