---
title: "Quick Demo"
author: "Lisa DeBruine & Daniel Lakens"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(faux)
```

Here is a quick example of setting up a simple study. 

## Set up an empty study object

```{r}
library(scienceverse)

mystudy <- study("IAT Demo")
```


## Add hypothesis 1

```{r}
mystudy <- add_hypothesis(
  study = mystudy,
  description = "Mean RT will be significantly slower in the incongruent condition compared to the congruent condition.",
  evaluation = "&", # all criteria must be true to accept
  id = "H1"
)
```


## Add analysis 1

Reference data you will attach later as ".data[id]" (here, we will use the id "iat").

```{r}
mystudy <- add_analysis(
  study = mystudy,
  func = "t.test",
  params = list(
    x = ".data[iat]$incongruent",
    y = ".data[iat]$congruent",
    alternative = "two.sided",
    paired = TRUE,
    var.equal = FALSE,
    conf.level = 0.95
  ),
  id = "H1_ttest"
)
```


## Add criteria to H1

Criteria results reference named list items returned from the analysis (e.g., the `t.test` in `H1_ttest` returns a list including `p.value` and `estimate`).

```{r}
# add criterion 1 to hypothesis 1
mystudy <- add_criterion(
  study = mystudy,
  result = "p.value",
  operator = "<",
  comparator = .05,
  hypothesis_id = "H1",
  analysis_id = "H1_ttest"
)

# add criterion 2 to hypothesis 1
mystudy <- add_criterion(
  study = mystudy,
  result = "estimate",
  operator = ">",
  comparator = 0,
  hypothesis_id = "H1",
  analysis_id = "H1_ttest"
)
```


## Add hypothesis 2

```{r}
mystudy <- add_hypothesis(
  study = mystudy,
  description = "Reaction times for congruent and incongruent trials will be signifiantly and positively correlated.",
  evaluation = "&",
  id = "H2"
)
```

## Add analysis 2

```{r}
mystudy <- add_analysis(
  study = mystudy,
  func = "cor.test",
  params = list(
    x = ".data[iat]$congruent",
    y = ".data[iat]$incongruent",
    alternative = "two.sided",
    method = "pearson",
    conf.level = 0.95
  ),
  id = "H2_cor"
)
```

## Add criteria to H2

```{r}
# add criterion 2 to hypothesis 2
mystudy <- add_criterion(
  study = mystudy,
  result = "p.value",
  operator = "<",
  comparator = 0.05,
  hypothesis_id = "H2",
  analysis_id = "H2_cor"
)

# add criterion 2 to hypothesis 2
mystudy <- add_criterion(
  study = mystudy,
  result = "estimate",
  operator = ">",
  comparator = 0,
  hypothesis_id = "H2",
  analysis_id = "H2_cor"
)
```

## Save the pre-registration archive file

```{r}
study_save(mystudy, "iat.json")
```

## Generate the pre-registration report

```{r, eval = FALSE}

study_report(mystudy, template="prereg", filename = "prereg")
```

```{r, results='asis', echo = FALSE}
# output section of the report
mystudy %>%
  output_hypotheses() %>%
  output_analyses()
```

## Load the study info from the archive file

```{r}
mystudy <- study("iat.json")
```


## Add data

Simulate some data using `faux::sim_design()` ([learn more about faux](https://debruine.github.io/faux/)).
```{r}
# simulate some data
dat <- faux::sim_design(
  within = list(condition = c("congruent", "incongruent")),
  n = 50, 
  mu = c(500, 750), 
  sd = 250, 
  r = .5
)
```

Add the data to the study object. Make sure the id matches the data id from the analysis parameters.
```{r}
mystudy <- add_data(mystudy, dat, id = "iat")
```

## Run analyses
```{r}
mystudy <- study_analyse(mystudy)
```


## Generate the post-registration report
```{r, eval = FALSE}

study_report(mystudy, template="postreg", filename = "postreg")
```

```{r, results='asis', echo = FALSE}
# output section of the report
mystudy %>%
  output_results()
```


## Piped setup

```{r}

mystudy <- study("IAT Demo")  %>%
  add_hypothesis(
    description = "Mean RT will be significantly slower in the incongruent condition compared to the congruent condition.",
    evaluation = "&", id = "H1") %>%
  add_analysis(
    func = "t.test",
    params = list(
      x = ".data[1]$incongruent",
      y = ".data[1]$congruent",
      alternative = "two.sided",
      paired = TRUE,
      var.equal = FALSE,
      conf.level = 0.95
    ),
    id = "H1_ttest"
  ) %>%
  add_criterion(
    result = "p.value",
    operator = "<",
    comparator = .05,
    hypothesis_id = "H1",
    analysis_id = "H1_ttest"
  ) %>%
  add_criterion(
    result = "estimate",
    operator = ">",
    comparator = 0,
    hypothesis_id = "H1",
    analysis_id = "H1_ttest"
  ) %>%
  add_hypothesis(
    description = "Reaction times for congruent and incongruent trials will be signifiantly and positively correlated.",
    evaluation = "&", id = "H2"
  ) %>%
  add_analysis(
    func = "cor.test",
    params = list(
      x = ".data[1]$congruent",
      y = ".data[1]$incongruent",
      alternative = "two.sided",
      method = "pearson",
      conf.level = 0.95
    ),
    id = "H2_cor"
  ) %>%
  add_criterion(
    result = "p.value",
    operator = "<",
    comparator = 0.05,
    hypothesis_id = "H2",
    analysis_id = "H2_cor"
  ) %>%
  add_criterion(
    result = "estimate",
    operator = ">",
    comparator = 0,
    hypothesis_id = "H2",
    analysis_id = "H2_cor"
  ) %>%
  add_data(dat) %>%
  study_analyse()
```

```{r, eval = FALSE}
mystudy %>%
  study_report(template="prereg", filename = "pre") %>%
  study_report(template="postreg", filename = "post")
```

