---
title: "Export Study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Export Study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
  .example {
    border: 2x solid grey;
    border-radius: 1em;
    padding: 0.5em;
  }
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(scienceverse)
# suppress most messages
scienceverse_options(verbose = FALSE)
```

## Create the meta-study file

Once you've created a scienceverse study, you can set up a project from the meta-study file. Here's a simple example testing group differences and simulating the data for the groups.

### Simulate Data

First, set up the basic study object and add a simulated dataset. 

```{r}
s <- study("Demo") %>%
  add_sim_data("dat", 
               within = list(time = c("morning", "night")),
               between = list(pet = c("dog", "cat")), 
               dv = c(y = "Happiness Score"),
               mu = list(dog = c(10, 10), 
                         cat = c(8)), 
               n = 30, r = 0.5, sd = 5, long = TRUE
  )
```

Plot the actual data and the design to check it looks like you expect.

```{r}
# plot faux data
dat <- get_data(s, "dat")
plot(dat)
# plot faux design specification
attr(dat, "design") %>% plot()
```

### Set up hypotheses

Here, we're adding two hypotheses, along with their associated analyses, evaluation criteria, and criteria for corroboration or falsification. These are very simple examples for demonstration, but you can include more complex analyses in curly brackets or as external .R files.

```{r}
s <- s %>%
  add_hypothesis("group", "Dogs will be happier than cats") %>%
  add_analysis("A1", scienceverse::t_test(y~pet, dat)) %>%
  add_criterion("p", "p.value", "<", 0.05) %>%
  add_criterion("dir", "estimate[1]", ">", "estimate[2]") %>%
  add_eval("c", "p & dir") %>%
  add_eval("f", "p & !dir") %>%
  
  add_hypothesis("time", "Pets will be happier in the morning than at night") %>%
  add_analysis("A2", scienceverse::t_test(y~time, dat)) %>%
  add_criterion("p", "p.value", "<", 0.05) %>%
  add_criterion("dir", "estimate[1]", ">", "estimate[2]") %>%
  add_eval("c", "p & dir") %>%
  add_eval("f", "p & !dir") %>%
  study_analyse()
```

NB: I'm using the scienceverse function `t_test()` instead of `t.test` because it gives you the Ns for each group in between-group comparisons (useful for meta-analyses) and also labels the estimates to avoid confusion.

Aftr you've run `study_analyse()`, printing the study object will show a summary of the evaluation.

```{r}
s
```

### Save and reload

You can save the study in JSON (machine-readable) format and reload it later. 

```{r}
study_save(s, "study.json")
remove(s)
s <- study("study.json")
```

## Get Results

You can get all the results with the `get_result()` function. If you don't specify the result name or the analysis ID, it defaults to all of the results of the first analysis. It returns a list that you can use, but will display an RMarkdown-formatted list if you print it (and set the chunk options to `results='asis'`).

```{r, results='asis'}
get_result(s, analysis_id = "A1")
```

```{r, results='asis'}
get_result(s, analysis_id = "A2")
```

Get a specific result from a specific analysis.

```{r}
get_result(study = s, 
           result = "p.value", 
           analysis_id = "A2", 
           digits = 3,
           return = "value")
```

Set `return` to "char" if you want to keep trailing zeros (this returns the number as a character string).

```{r}
get_result(study = s, 
           result = "parameter", 
           analysis_id = 1, 
           digits = 5,
           return = "char")
```

### Linked values

You can display the value as a link if you set `return` to "html". You can use the shorthand function `get_html` if you only have one study object loaded. The digits default to the global option, so you can set that as shown.
 
```{r}
options(digits = 3)
get_html("p.value")
```

You'll probably want to use `get_html()` inline most of the time like below. The links go to a section at the end of this document that is created with [`make_script()`](#Appendix).

<div class="example">
For hypothesis 1, dogs were happier than cats 
  (t = `r get_html("statistic", 1, 2)`, 
  df = `r get_html("parameter", 1, 1)`, 
  p = `r get_html("p.value", 1, 3)`,
  95% CI = [`r get_html("conf.int[1]", 1, 2)`, 
             `r get_html("conf.int[2]", 1, 2)`]).
  
For hypothesis 2, the subjects were not happier in the morning than at night 
  (t = `r get_html("statistic", 2, 2)`, 
  df = `r get_html("parameter", 2, 1)`, 
  p = `r get_html("p.value", 2, 3)`,
  95% CI = [`r get_html("conf.int[1]", 2, 2)`, 
             `r get_html("conf.int[2]", 2, 2)`]).
</div>


### External analysis file

You can also display the value as a link to an external analysis file created with `make_script`. For the shorthand function, set the analysis link with `scienceverse_options()` as shown.

```{r, message=FALSE}
# create the analysis .Rmd and .html file, 
# or you can do this outside this script
make_script(s, "analysis.Rmd")
rmarkdown::render("analysis.Rmd", quiet = TRUE) 

scienceverse_options(analysis_link = "analysis.html")
get_html("p.value")
```

Here the inline links now go to an external file:

<div class="example">
For hypothesis 1, dogs were happier than cats 
  (t = `r get_html("statistic", 1, 2)`, 
  df = `r get_html("parameter", 1, 1)`, 
  p = `r get_html("p.value", 1, 3)`,
  95% CI = [`r get_html("conf.int[1]", 1, 2)`, 
             `r get_html("conf.int[2]", 1, 2)`]).
  
For hypothesis 2, the subjects were not happier in the morning than at night 
  (t = `r get_html("statistic", 2, 2)`, 
  df = `r get_html("parameter", 2, 1)`, 
  p = `r get_html("p.value", 2, 3)`,
  95% CI = [`r get_html("conf.int[1]", 2, 2)`, 
             `r get_html("conf.int[2]", 2, 2)`]).
</div>


## Inline Script {#Appendix}

Add the script inline at the end with the function [`make_script`](../reference/make_script.html). Set `header` to `FALSE` to omit the YAML header. Set `header_lvl` to change the default starting header level of 2.This function will save the data and codebooks in a folder called "data". Set the argument `data_path` to NULL to include the actual data in the text of the script. For large datasets, you'll want to leave it as the default "data" folder (or set a custom folder name).

```{r, results='asis'}
make_script(s, header_lvl = 3, header = FALSE) %>% cat()
```

```{r, include= FALSE}
# clean up data directory for vignette-creation purposes
unlink("data", recursive = TRUE)
remove("analysis.html")
```



