---
title: "Better Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Better Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(scienceverse)
```

Scienceverse provides some functions to replace common tests like `t.test` to provide data that is useful for meta-analyses. There is only one very basic example right now, but we will expand this in the future.

## t.test vs t_test

The output of `t.test` does not return the number of data points per group. You can figure this out from the df for one-sample and paired t-tests, but not for two-sample t-tests. The function `t_test` returns a list with the same items as `t.test` plus `n`, which is the number of data points or pairs included in the one-sample or paired t-test (after excluding NAs, which `t.test` does silently), or a 2-number named vector of the number of data points in each group for a two-sample t-test.

### Data

First, we'll set up an example data frame with 20 rows, two continuous columns (x and y) and a grouping column (g).

```{r}
dat <- data.frame(
  x = rnorm(20),
  y = rnorm(20, 0.5),
  g = rep(c("G1", "G2"), 10)
)
```

### One sample

```{r}
res <- t_test(dat$x)
res$n
```

### Paired samples

```{r}
res <- t_test(dat$x, dat$y, paired = TRUE)
res$n
```

### Independent samples

```{r}
Grp1 <- dat[dat$g== "G1", ]$x
Grp2 <- dat[dat$g== "G2", ]$x
res <- t_test(Grp1, Grp2)
res$n
```

The `n` is a named vector for independent-samples t-tests. The name comes from the variable names of `x` and `y` (not their original names in the data frame), unless you specify it directly with the `names` argument.

```{r}
res <- t_test(Grp1, Grp2, names = c("A", "B"))
res$n
```

You can also use the formula version.

```{r}
res <- t_test(x ~ g, data = dat)
res$n
```
