---
title: "Make Study from Grobid XML"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{grobid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(scienceverse)
library(dplyr)
```

## Load from XML

The function `study_from_xml()` can read XML files in [TEI](https://tei-c.org/) format created by [grobid](https://grobid.readthedocs.io/). 

```{r}
grobid_dir <- system.file("grobid", package="scienceverse")
filename <- file.path(grobid_dir, "incest.pdf.tei.xml")
s <- study_from_xml(filename)
```

## Full Text

You can access a parsed table of the full text of the paper via `s$full_text`

```{r}
count(s$full_text, section_class)
```

However, you may find it more convenient to use the function `search_full_text()`.

```{r}
search_full_text(s, "significan[t|ce]")
```

(Note that grobid is often bad at parsing sentences with statistics, and will break them into multiple sentences. We are working on ways of detecting and fixing this.)

References are omitted by default, but you can add them back in.

```{r}
search_full_text(s, "third-party", section = "intro", refs = TRUE)
```


## Batch Processing

The function `study_from_xml()` also works on a folder of XML files, returning a list of scienceverse study objects, and `search_full_text()` works on such a list.

```{r}
studies <- study_from_xml(grobid_dir)

search_full_text(studies, "hypothesi", section = "intro")
```

