---
title: "Power Demo (ANOVA)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Power Demo (ANOVA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

How do you test a specific hypothesis for a complex hypothesis, such as a 4x3x2 mixed-design ANOVA? For example, suppose you have a hypothesis about two within-subject factors (A and B) in three between-subject conditions (control, C1 and C2). You predict that there will be no effects of A or B (and no interaction) in the control condition, an effect of B only on A1 in the first experimental condition, and an effect of B only on A2 in the second experimental condition. 

The point of this tutorial is to show just how complex the decision criteria can be for an interaction. I don't think I've chosen ideal criteria, but this is a model for how to start thinking about how you'd specify hypothesis corroboration criteria that are more sophisticated than just "predict a significant 3-way interaction".

## Packages Used

```{r}
suppressPackageStartupMessages({
  library(scienceverse)
  library(faux)
  library(afex)
  library(tidyverse)
})
scienceverse_options(verbose = FALSE)
```


## Test Data Simulation

First, simulate some data so you can build your analysis structure. I've built in a pretty big interaction, with nothing at all going on in the control condition, and opposite interactions among the within-subject factors in experimental conditions 1 and 2. There are 20 observations (`n`) in each between cell. Check the 

```{r}
dat <- faux::sim_design(
   within = list(A = c("A1", "A2", "A3", "A4"),
                 B = c("B1", "B2")),
   between = list(C = c("C0", "C1", "C2")),
   n = 20,
        # A1_B1, A1_B2, A2_B1 ... A4_B2
   mu = c(0, 0,  0,  0,  0,  0,  0,  0,  # C0
          0, 0, .2,  0, .4,  0, .6,  0,  # C1
          0, 0,  0, .2,  0, .4,  0, .6), # C2
   sd = 1,
   r = 0.5,
   id = "sub_id",
   dv = "score",
   long = TRUE)
```

## Test analysis

I'm using the `afex::aov_ez()` function to run an ANOVA. This maps on pretty straightforwardly to the way you simulate the data in {faux}.

```{r}
# run ANOVA
omnibus <- afex::aov_ez(id = "sub_id",
                        dv = "score",
                        between = "C",
                        within = c("A", "B"),
                        data = dat,
                        return = "afex_aov")
```

{scienceverse} requires any numbers you use from an analysis to be returned as a named list, and we only care about the 3-way interaction for this study, so we'll use {dplyr} and {tidyr} to get just the values we want.

```{r}
# get stats for 3-way interaction
ixn <- anova(omnibus)["C:A:B", ]

# calculate group means
means <- group_by(dat, A, B, C) %>%
  summarise(m = mean(score), .groups = "drop") %>%
  tidyr::unite("grp", A, B, C)

m <- means$m
names(m) <- means$grp

list(
  p = ixn$`Pr(>F)`,
  F = ixn$F,
  means = m
) %>% str()
```

Now you might also want to do some analyses to check that the pattern of any significant 3-way interaction is consistent with your hypothesis.

```{r}
aov_control <- afex::aov_ez(id = "sub_id",
                        dv = "score",
                        within = c("A", "B"),
                        data = filter(dat, C == "C0"),
                        return = "afex_aov")

aov_c1 <- afex::aov_ez(id = "sub_id",
                        dv = "score",
                        within = c("A", "B"),
                        data = filter(dat, C == "C1"),
                        return = "afex_aov")

aov_c2 <- afex::aov_ez(id = "sub_id",
                        dv = "score",
                        within = c("A", "B"),
                        data = filter(dat, C == "C2"),
                        return = "afex_aov")

list(
  control = anova(aov_control)["A:B", "Pr(>F)"],
  cond1 = anova(aov_c1)["A:B", "Pr(>F)"],
  cond2 = anova(aov_c2)["A:B", "Pr(>F)"]
) %>% str()
```


## Set up Scienceverse study

Set up the study by adding simulated data (use the argument from `faux::sim_design()` above), the analyses above (make sure to add the library calls at the start of the analysis code or preface each function with the package name).

```{r}
s <- study("Demo") %>%
  
  add_sim_data(data_id = "dat",
    within = list(A = c("A1", "A2", "A3", "A4"),
                  B = c("B1", "B2")),
    between = list(C = c("C0", "C1", "C2")),
    n = 20,
         # A1_B1, A1_B2, A2_B1 ... A4_B2
    mu = c(0, 0,  0,  0,  0,  0,  0,  0,  # C0
           0, 0, .2,  0, .4,  0, .6,  0,  # C1
           0, 0,  0, .2,  0, .4,  0, .6), # C2
    sd = 1,
    r = 0.5,
    id = "sub_id",
    dv = "score",
    long = TRUE) %>%
  
  add_analysis("ANOVA", code = {
    library(dplyr) # add for pipes
    
    # run ANOVA
    omnibus <- afex::aov_ez(id = "sub_id",
                            dv = "score",
                            between = "C",
                            within = c("A", "B"),
                            data = dat,
                            return = "afex_aov")

    # get stats for 3-way interaction
    ixn <- anova(omnibus)["C:A:B", ]

    # calculate group means
    means <- group_by(dat, A, B, C) %>%
      summarise(m = mean(score), .groups = "drop") %>%
      tidyr::unite("grp", A, B, C)
    
    m <- means$m
    names(m) <- means$grp
    
    list(
      p = ixn$`Pr(>F)`,
      F = ixn$F,
      means = m
    )
  })
```

Check the analysis works.

```{r}
study_analyse(s) %>% get_result()
```

Add the extra analyses. They can go in the code of the main analysis and their results can be added to the returned list, or they can be added as separate analysis objects.

```{r}
s <- s %>%
  add_analysis("C0", code = {
    # run ANOVA
    aov_c <- afex::aov_ez(
      id = "sub_id",
      dv = "score",
      within = c("A", "B"),
      data = dplyr::filter(dat, C == "C0"),
      return = "afex_aov")

    # get stats for 2-way interaction
    ixn <- anova(aov_c)["A:B", ]

    list(
      p = ixn$`Pr(>F)`,
      F = ixn$F
    )
  }) %>%
  
  add_analysis("C1", code = {
    # run ANOVA
    aov_c <- afex::aov_ez(
      id = "sub_id",
      dv = "score",
      within = c("A", "B"),
      data = dplyr::filter(dat, C == "C1"),
      return = "afex_aov")

    # get stats for 2-way interaction
    ixn <- anova(aov_c)["A:B", ]

    list(
      p = ixn$`Pr(>F)`,
      F = ixn$F
    )
  }) %>%
  
  add_analysis("C2", code = {
    # run ANOVA
    aov_c <- afex::aov_ez(
      id = "sub_id",
      dv = "score",
      within = c("A", "B"),
      data = dplyr::filter(dat, C == "C2"),
      return = "afex_aov")

    # get stats for 2-way interaction
    ixn <- anova(aov_c)["A:B", ]

    list(
      p = ixn$`Pr(>F)`,
      F = ixn$F
    )
  })

```

Check the new analyses works.

```{r}
s <- study_analyse(s) 

get_result(s, analysis_id = "C0")
get_result(s, analysis_id = "C1")
get_result(s, analysis_id = "C2")
```


Add a hypothesis, specify criteria, and add evaluation rules You can add rules for both "corroboration" and "falsification", but here we just have corroboration rules.

```{r}
  
s <- s %>%
  add_hypothesis("IXN", "There will be a significant 3-way interaction among factors A, B and C") %>%
  # C:A:B interaction significant
  add_criterion("sig_ixn", 
                result = "p", 
                operator = "<", 
                comparator = 0.05,
                analysis_id = "ANOVA") %>%

  # add criteria to show the pattern is as expected
  
  # A1 < A2 < A3 < A4 from B1 in C1
  add_criterion("c1_1",
                result = "mean$A1_B1_C1", 
                operator = "<", 
                comparator = "mean$A2_B1_C1",
                analysis_id = "ANOVA") %>%
  
  add_criterion("c1_2",
                result = "mean$A2_B1_C1", 
                operator = "<", 
                comparator = "mean$A3_B1_C1",
                analysis_id = "ANOVA") %>%
  
  add_criterion("c1_3",
                result = "mean$A3_B1_C1", 
                operator = "<", 
                comparator = "mean$A4_B1_C1",
                analysis_id = "ANOVA") %>%
  
  # A1 < A2 < A3 < A4 from B2 in C2
  add_criterion("c2_1",
                result = "mean$A1_B2_C2", 
                operator = "<", 
                comparator = "mean$A2_B2_C2",
                analysis_id = "ANOVA") %>%
  
  add_criterion("c2_2",
                result = "mean$A2_B2_C2", 
                operator = "<", 
                comparator = "mean$A3_B2_C2",
                analysis_id = "ANOVA") %>%
  
  add_criterion("c2_3",
                result = "mean$A3_B2_C2", 
                operator = "<", 
                comparator = "mean$A4_B2_C2",
                analysis_id = "ANOVA") %>%
  
  # all criteria must be true to corroborate
  add_eval(type = "corroboration",
          evaluation = "sig_ixn & c1_1 & c1_2 & c1_3 & c2_1 & c2_2 & c2_3")
```


## Run power simulation

Running `study_power()` returns a study object that has `rep` values stored for the analysis criteria and evaluations. `get_power()` returns the evaluation for each hypothesis and, optionally, the results values for each analysis.

```{r}
s <- study_power(s, rep = 100)

power <- get_power(s, values = TRUE)

```


```{r}
str(power$power)
```


## Plot the resulting values

```{r, fig.width = 9, fig.height = 4, out.width = "100%"}

data.frame(
  ixn = power$results$ANOVA$p,
  c0 = power$results$C0$p,
  c1 = power$results$C1$p,
  c2 = power$results$C2$p
) %>%
  gather("analysis", "p", ixn:c2) %>%
  ggplot(aes(x = p, fill = analysis)) + 
  geom_histogram(binwidth = 0.05, color = "black", 
                 boundary = 0, show.legend = FALSE) +
  facet_grid(~analysis) +
  xlab("p-values")
```


```{r, fig.width = 5, fig.height = 5, out.width = "100%"}

t(power$results$ANOVA$means) %>%
  as.data.frame() %>%
  pivot_longer(
    cols = A1_B1_C0:A4_B2_C2,
    names_to = c("A", "B", "C"),
    names_sep = "_",
    values_to = "mean"
  ) %>%
  ggplot(aes(x = A, y = mean, colour = B)) + 
  geom_violin(alpha = 0.5, position = position_dodge(width = 0)) +
  facet_grid(C~.)
```


```{r, fig.width = 8, fig.height = 5, out.width = "100%"}

t(power$results$ANOVA$means) %>%
  as.data.frame() %>%
  mutate(rep = row_number()) %>%
  pivot_longer(
    cols = A1_B1_C0:A4_B2_C2,
    names_to = c("A", "B", "C"),
    names_sep = "_",
    values_to = "mean"
  ) %>%
  pivot_wider(
    names_from = B,
    values_from = `mean`
  ) %>%
  mutate(B_diff = B2 - B1) %>%
  ggplot(aes(x = A, y = B_diff, colour = C)) + 
  geom_violin(alpha = 0, position = position_dodge(width = 0))
```

